---
title: "RDS Catalog Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rds.catalog}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
   table {
    width: 100%;
  }
</style>

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library("rds.r")
library("knitr")
```

## Browsing a Catalog

Catalogs in RDS are simply collections of data products. A data product is the object in RDS that holds data and any documentation or metadata that surrounds the data. The catalog contains a list of data products and we can get all of them or a single product if desired.

```{r}
# set up the host / domain name of the RDS instance
host = "http://dev.richdataservices.com"

# connect to the server
rds <- get.rds(host)

# get a catalog
census <- rds.r::getCatalog(rds, "uscensus")
```


### Get Specific Data Product
We can get the 2010 Census Household / Person data product and display some of its information in the following section

```{r}
dataProduct <- rds.r::getDataProduct(census, "census2010-hp")
```

### `r dataProduct@name`



`r dataProduct@description`