---
title: "RDS Server Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RDS Server Examples}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

<style>
   table {
    width: 100%;
  }
</style>

## Browsing a Catalog

Catalogs in RDS are simply collections of data products. A data product is the object in RDS that holds data and any documentation or metadata that surrounds the data. The catalog contains a list of data products and we can get all of them or a single product if desired.

## Setup

First we will set up the RDS server connection and get a catalog

```{r, message=FALSE, warning=FALSE}
library("rds.r")
library("knitr")

# set up the host / domain name of the RDS instance
host = "192.168.1.3"

# If needed you can also specify an optional port that the instance is running on
port = 8080

# connect to the server
rds <- get.rds(host, port = port)

# get a catalog
anes <- rds.r::getCatalog(rds, "anes");
```


## Get all Data Products

```{r, message=FALSE, warning=FALSE}
dataProducts <- rds.r::getDataProducts(anes)
```

### Get Specific Data Product

```{r}
anes1948 <- rds.r::getDataProduct(anes, "anes1948")
```